---
title: "Depression Among Adults"
format: html
author: "Mariama Marish Keita"
date: "2025-02-15"
---

## Project Description

This dataset contains information on adult depression among males and females from different ethnicities in the USA.

## Research questions

- Which ethnic group has the highest prevalence of depression over the years?

- Is there a significant difference in depression rates based on income levels
Statistical Association

- Is there a correlation between weighted frequency and percentage of depression?

- Are there significant differences in depression rates between different age groups ?

- Show the distribution of depression rates across different groups.


## Loading of Libraries 

For the library installation i prefer to use pacman package . It is a package manager that manages other packages. 


```{r setup, include=FALSE}
if (!require(readr)) install.packages("readr") or 

# if(!require(pacman)) install.packages("pacman")
# pacman::pload(readr, ggplot2, tidyr,dplyr, nortest, magrittr)
library(readr) # It is used for reading and writing data efficiently
library(ggplot2) # Used for data visualization
library(tidyr) # this pacakge is use for data cleaning and reshaping 
library(dplyr) # They are used for data manipulation and transformation
library(nortest) # Provides multiple statistical tests to check if a dataset follows a normal distribution
library(magrittr) # used for writting cleaner, more readable code by providing the pipe operator.
```

## Import dataset

The `dirname()` function in R extracts the directory (folder ) path from a given file path. It is useful when working with file paths, directories. 

```{r}
#Set working directory

# the followings are not needed - commented them out
# setwd("/Users/admin/Desktop/adult-depression-lghc-indicator-24.csv")
# getwd()

# create a file_path for easy access
file_path <- file.path(dirname(getwd()), "depression_data", "adult-depression-lghc-indicator-24.csv")

# import the data
depression_data <- read.csv(file_path)
```

## Description of the dataset 

This dataset contains information on adult depression among males and females from different ethnicities in the USA.

This dataset includes `r nrow(depression_data)` rows and `r ncol(depression_data)` columns, covering data from 2012 to 2018. 


The column `r colnames(depression_data)[1]` represents the year in which the data was recorded.  

## or we can also say The column `r colnames(depression_data)[which(colnames(depression_data) == "Year)]`

- `r colnames(depression_data)[2]` represents a category used to divide the data (eg race, age,income). 
- `r colnames(depression_data)[3]` specifies the subgroup within the Strata.  
- `r colnames(depression_data)[4]` represents the count of individuals in the given category.  
- `r colnames (depression_data)[5]`  provides a weighted estimate of the number of people in the population.  
- `r colnames(depression_data)[6] ` represents the percentage of the population experiencing depression condition.  
- `r colnames(depression_data)[7]` represents the lower bound of the 95% confidence interval.  
- `r colnames(depression_data)[8]` represents the upper bound of the 95% confidence interval.

## Data analysis 

```{r}
# filter out rows with Sex and Total in the 'Strata' column
data_1 <- depression_data %>%
  dplyr::filter(!(Strata %in% c("Total", "Sex")))

# plot the depression level distribution across the different strata over years
p <- ggplot(data_1, aes(fill = Strata.Name, x = Strata, y = Percent)) +
  geom_bar(position = "dodge", stat = "identity") +
  facet_wrap(~Year)
```

```{r}
ggplot(depression_data,
       aes(x = Year, y = Percent, group = Strata.Name,
           color = Strata.Name)) +
  geom_line() + 
  facet_wrap(~Strata)
```


- are there  significant difference in depression rates between males and females over the years?

## statistical tests

To know if there is a statistical difference in depression btw male and female we first need to check if the distribution is normal then we hypothesised that there is a statistical difference between two sexes. 

we used the shapiro to confirm that. 
```{r}
data_2 <- depression %>%
  dplyr::filter(!(Strata %in% c("Total")))

shapiro.test(data_2$Frequency)

# since p -value is less that 0.03 then the distribution is not  normal 


# ops i have to come back and work on this 
```
