---
title: "Depression Among Adults"
format: html
author: "Mariama Marish Keita"
date: "2025-02-15"
---

## Project Description

Provide the project description here.

## Loading of Libraries 

* Provide instructions for library installation, and
* reasons why you are using every library.

```{r setup, include=FALSE}
library(readr)
library(ggplot2)
library(tidyr)
library(dplyr)
library(nortest)
library(magrittr)
```

## Import dataset

```{r}
#Set working directory

setwd("/Users/admin/Desktop/adult-depression-lghc-indicator-24.csv")
getwd()
# create a file_path for easy access

file_path <- file.path(getwd(), "adult-depression-lghc-indicator-24.csv")

depression_data <- read.csv(file_path)
```

## Description of the dataset 


This dataset contains information on adult depression among males and females from different ethnicities in the USA. It includes 161 rows and 8 columns, covering data from 2012 to 2018.  

- Column 1 represents the year in which the data ws recorded.  
- Column 2 represents a category used to divide the data (eg race, age,income). 
- Column 3 specifies the subgroup within the Strata.  
- Column 4 represents the count of individuals in the given category.  
- Column 5 provides a weighted estimate of the number of people in the population.  
- Column 6 represents the percentage of the population experiencing the condition.  
- Column 7 represents the lower bound of the 95% confidence interval.  
- Column 8 represen the upper bound of the 95% confidence interval.

# Research Questions 

- How has the prevalence of adult depression changed from 2012 to 2018?
```{r}
ggplot(depression_data, aes(x = Year, y = Percent, group = Strata, color = Strata)) +
  geom_line() + 
  geom_point() +
  labs(title = "Trends in Adult Depression (2012-2018)",
       x = "Year", y = "Prevalence (%)")

```


- are there  significant difference in depression rates between males and females over the years?

```{r}


sex_depression <- depression_data %>%
  filter(Strata == "Sex") %>%
  ggplot(aes(x = Strata.Name, y = Percent, fill = Strata.Name)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Depression Rates in Males and Females Over the Years",
       x = "Sex", y = "Prevalence (%)") +
  theme_minimal()


```


- How does race/ethnicity influence depression rates?
```{r}
ggplot(depression_data, aes(x = Strata.Name, y = Percent, fill = Strata.Name)) +
  geom_bar(stat = "identity") +
  labs(title = "Depression Rates by Race/Ethnicity",
       x = "Race/Ethnicity", y = "Prevalence (%)")

```


- Which ethnic group has the highest prevalence of depression over the years?

- Is there a significant difference in depression rates based on income levels
Statistical Associations

- Is there a correlation between weighted frequency and percentage of depression?

- Are there significant differences in depression rates between different age groups ?

- Show the distribution of depression rates across different groups.

```{r}
ggplot(depression_data, aes(x = Strata.Name, y = Percent, fill = Strata.Name)) +
  geom_boxplot() +
  labs(title = "Distribution of Depression Rates by Group",
       x = "Group", y = "Prevalence (%)")

```
is the distribution normal for age ?

```{r}
ggplot(depression_data, aes(x = Age)) +
  geom_histogram(aes(y = ..density..), bins = 30, fill = "blue", alpha = 0.5) +
  geom_density(color = "pink") +
  labs(title = "Distribution of Age", x = "Age", y = "Density")
# or 

depression_data %>%
  group_by(Year, Strata.Name) %>%
  summarise(p_value = shapiro.test(Age)$p.value)


```
