---
title: "Depression Among Adults"
format: html
author: "Mariama Marish Keita"
date: "2025-02-15"
---

## Project Description

This project analyzes data collected from individuals in the United States who experienced depression between `2012` and `2017`. The dataset includes individuals from diverse backgrounds, categorized by race, gender, income, education level, and age.

The racial categories include Black, White, Hispanic, Asian/Islander, and Other. Gender is classified as male or female. Educational background is divided into five groups: no high school diploma, high school graduate, some college or technical school, college graduate, and postgraduate.

The study also examines whether depression is influenced by age, with age groups categorized as 18–34, 35–44, 45–54, 55–64, and 65 and above. Additionally, it explores the potential correlation between depression and income levels, with income brackets of less than $20,000, $20,000–$34,999, $35,000–$49,999, $50,000–$74,900, $75,000–$99,999, and $100,000 or more per year.

The primary objective of this project is to determine whether there is a correlation between depression and factors such as gender, age, income, and education. It specifically aims to identify whether depression is more prevalent among certain groups, such as males or females, different age ranges, or individuals with varying income levels and educational backgrounds.


## Research questions

- Which ethnic group has the highest prevalence of depression over the years?

- Is there a significant difference in depression rates based on income levels
Statistical Association

- Is there a correlation between weighted frequency and percentage of depression?

- Are there significant differences in depression rates between different age groups ?



## Loading of Libraries 

For the library installation I prefer to use the {pacman} package . It is a package manager that manages and facilitate the installation of other packages. 


```{r setup, include=FALSE}


# if(!require(pacman)) install.packages("pacman")
# pacman::pload(readr, ggplot2, tidyr,dplyr, nortest, magrittr)
library(readr) # It is used for reading and writing data efficiently
library(ggplot2) # Used for data visualization
library(tidyr) # this package is use for data cleaning and reshaping 
library(dplyr) # They are used for data manipulation and transformation
library(nortest) # Provides multiple statistical tests to check if a dataset follows a normal distribution
library(magrittr) # used for writting cleaner, more readable code by providing the pipe operator.
```

## Import dataset

The `dirname()` function in R extracts the directory (folder ) path from a given file path. It is useful when working with file paths, directories. 

```{r}
#Set working directory

# the followings are not needed - commented them out
# setwd("/Users/admin/Desktop/adult-depression-lghc-indicator-24.csv")
# getwd()

# create a file_path for easy access
file_path <- file.path(dirname(getwd()), "adult-depression-lghc-indicator-24.csv")

# import the data
depression_data <- read.csv(file_path)
```

## Description of the dataset 

This dataset contains information on adult depression among males and females from different ethnicities in the USA.

This dataset includes `r nrow(depression_data)` rows and `r ncol(depression_data)` columns, covering data from 2012 to 2018. 


The column `r colnames(depression_data)[1]` represents the year in which the data was recorded.  

## or we can also say The column `r colnames(depression_data)[which(colnames(depression_data) == "Year)]`

- `r colnames(depression_data)[2]` represents a category used to divide the data (eg race, age,income). 
- `r colnames(depression_data)[3]` specifies the subgroup within the Strata.  
- `r colnames(depression_data)[4]` represents the count of individuals in the given category.  
- `r colnames (depression_data)[5]`  provides a weighted estimate of the number of people in the population.  
- `r colnames(depression_data)[6] ` represents the percentage of the population experiencing depression condition.  
- `r colnames(depression_data)[7]` represents the lower bound of the 95% confidence interval.  
- `r colnames(depression_data)[8]` represents the upper bound of the 95% confidence interval.

## Data analysis 

```{r}
# filter out rows with Sex and Total in the 'Strata' column
data_1 <- depression_data %>%
  dplyr::filter(!(Strata %in% c("Total", "Sex")))

# plot the depression level distribution across the different strata over years
p <- ggplot(data_1, aes(fill = Strata.Name, x = Strata, y = Percent)) +
  geom_bar(position = "dodge", stat = "identity") +
  facet_wrap(~Year)
```

```{r}
ggplot(depression_data,
       aes(x = Year, y = Percent, group = Strata.Name,
           color = Strata.Name)) +
  geom_line() + 
  facet_wrap(~Strata)
```


- are there  significant difference in depression rates between males and females over the years?

## statistical tests

To know if there is a statistical difference in depression btw male and female we first need to check if the distribution is normal then we hypothesised that there is a statistical difference between two sexes. 

we used the shapiro to confirm that. 
```{r}
data_2 <- depression %>%
  dplyr::filter(!(Strata %in% c("Total")))

shapiro.test(data_2$Frequency)

# since p -value is less that 0.03 then the distribution is not  normal 


```
